---
title: 'Gov 2018: Lab 5 Random Forests'
author:
- 'Your name: '
date: 'February 22, 2022'
output:
  pdf_document: default
  html_document: default
---

This exercise is based off of Muchlinski, David, David Siroky, Jingrui He, and Matthew Kocher. 2016. "Comparing Random Forest with Logistic Regression for Predicting Class-Imbalanced Civil War Onset Data". \emph{Political Analysis}.


Descriptions of the relevant variables in the data file `data_full.rds` are:

 Name                             Description
 -------------------------------- ----------------------------------------------------------
 `warstds`                        Factor, `peace` and `war`
 `year`                           Numeric for year of obs
 
 And a list of 90 covariates from the Sambanis dataset: "ager", "agexp", "anoc", "army85", "autch98", "auto4", "autonomy", "avgnabo", "centpol3", "coldwar", "decade1", "decade2", "decade3", "decade4", "dem", "dem4", "demch98", "dlang", "drel", "durable", "ef", "ef2", "ehet", "elfo", "elfo2", "etdo4590", "expgdp", "exrec", "fedpol3", "fuelexp", "gdpgrowth", "geo1", "geo2", "geo34", "geo57", "geo69", "geo8", "illiteracy", "incumb", "infant", "inst", "inst3", "life", "lmtnest", "ln_gdpen", "lpopns", "major", "manuexp", "milper", "mirps0", "mirps1", "mirps2", "mirps3", "nat_war", "ncontig", "nmgdp", "nmdp4_alt", "numlang", "nwstate", "oil", "p4mchg", "parcomp", "parreg", "part", "partfree", "plural", "plurrel", "pol4", "pol4m", "pol4sq", "polch98", "polcomp", "popdense", "presi", "pri", "proxregc", "ptime", "reg", "regd4_alt", "relfrac", "seceduc", "second", "semipol3", "sip2", "sxpnew", "sxpsq", "tnatwar", "trade", "warhist", "xconst".

```{r setup, include=FALSE}
# Colorblind friendly palette with grey:
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
library(randomForest) #for random forests
library(caret) # for CV folds and data splitting
library(ROCR) # for diagnostics and ROC plots/stats
library(pROC) # same as ROCR
library(stepPlr) # Firth's rare events logistic regression ("plr") implemented by caret
library(doMC) # for using multiple processor cores
library(xtable) # for writing Table 1 in Latex
library(tidyverse)

#Read in rds data
data.full<-readRDS("data_full.rds")
# str(data.full)
#Since war is a rare event, Firth's rare event logistic regression (or penalization logistic regression) has been often used in the literature
table(data.full$warstds) 

#Distribute workload over multiple cores for faster computation
detectCores()
# registerDoMC(cores=6) # specify the number of cores to use for parallel execution
set.seed(666) # random seed number from the replication codes
```

We're going to use the cross-validation function from the `caret` package. Set aside the years 1999 and 2000 for testing data.

```{r}
# caret::trainControl() controls parameters for train
tc<-caret::trainControl(method="cv", # the resampling method
                        number=10, # the number of folds 
                        summaryFunction=twoClassSummary, # a function to compute performance metrics across resamples. 
                                                         # twoClassSummary computes sensitivity, specificity and the AUC
                        classProb=T, # class probabilities be computed for classification models 
                                     # (along with predicted values) in each resample
                        savePredictions = T)

# Set train data
data.train<-subset(data.full,year<1999)
```

## Question 1

We're going to compare several model specifications using classic/penalized logistic regressions with a random forest model.

(a) The Fearon & Laitin model (2003) "FL" can be described as the following:

`as.factor(warstds) ~ warhist + ln_gdpen + lpopns + lmtnest + ncontig + oil + nwstate + inst3 + pol4 + ef + relfrac`

Please run the `train` function in the `caret` library with metric as `ROC`, method as `glm`, family as `binomial` (FL used logistic), trControl as our set `tc`, and your training data, on the above specification. Do the same for a penalized logistic regression (`method="plr"`).

```{r}

# Logistic regression
FL_mod <- train(as.factor(warstds) ~ warhist + ln_gdpen + lpopns + lmtnest + ncontig + oil + nwstate + inst3 + pol4 + ef + relfrac, data.train, 
      method = "glm", metric = "ROC", family = binomial(link = "logit"), trControl = tc) 

# Penalized logistic regression
FL_mod_plr <- train(as.factor(warstds) ~ warhist + ln_gdpen + lpopns + lmtnest + ncontig + oil + nwstate + inst3 + pol4 + ef + relfrac, data.train, 
      method = "plr", metric = "ROC", trControl = tc) 

```


(b) The Collier & Hoeffler model (2004) (CH) can be described as the following:

`as.factor(warstds) ~ sxpnew + sxpsq + ln_gdpen + gdpgrowth + warhist + lmtnest + ef + popdense + lpopns + coldwar + seceduc + ptime`

Please run the `train` function in the `caret` library with metric as `ROC`, method as `glm`, family as `binomial` (CH used logistic), trControl as our set `tc`, and your training data, on the above specification. Do the same for a penalized logistic regression (`method="plr"`).

```{r}

# Logistic regression
CH_mod <- train(as.factor(warstds) ~ sxpnew + sxpsq + ln_gdpen + gdpgrowth + warhist + lmtnest + ef + popdense + lpopns + coldwar + seceduc + ptime, data.train, 
      method = "glm", metric = "ROC", family = binomial(link = "logit"), trControl = tc) 

# Penalized Logistic regression
CH_mod_plr <- train(as.factor(warstds) ~ sxpnew + sxpsq + ln_gdpen + gdpgrowth + warhist + lmtnest + ef + popdense + lpopns + coldwar + seceduc + ptime, data.train, 
      method = "plr", metric = "ROC", trControl = tc)

```


(c) The Hegre & Sambanis model (2006) (HS) can be described as the following:

`as.factor(warstds) ~ lpopns + ln_gdpen + inst3 + parreg + geo34 + proxregc + gdpgrowth + anoc +  partfree + nat_war + lmtnest + decade1 + pol4sq + nwstate + regd4_alt + etdo4590 + milper +  geo1 + tnatwar + presi`

Please run the `train` function in the `caret` library with metric as `ROC`, method as `glm`, family as `binomial` (HS used logistic), trControl as our set `tc`, and your training data, on the above specification. Do the same for a penalized logistic regression (`method="plr"`).

```{r}

# Logistic regression
HS_mod <- train(as.factor(warstds) ~ lpopns + ln_gdpen + inst3 + parreg + geo34 + proxregc + gdpgrowth + anoc +  partfree + nat_war + lmtnest + decade1 + pol4sq + nwstate + regd4_alt + etdo4590 + milper +  geo1 + tnatwar + presi, data.train, 
      method = "glm", metric = "ROC", family = binomial(link = "logit"), trControl = tc)

# Penalized Logistic regression
HS_mod_plr <- train(as.factor(warstds) ~ lpopns + ln_gdpen + inst3 + parreg + geo34 + proxregc + gdpgrowth + anoc +  partfree + nat_war + lmtnest + decade1 + pol4sq + nwstate + regd4_alt + etdo4590 + milper +  geo1 + tnatwar + presi, data.train, 
      method = "plr", metric = "ROC", trControl = tc)

```


(d) Finally, run a random forest model on the outcome `warstds` with all regressors (except `year`) using `train`, metric as `ROC`, sampsize as `c(30,90)`, importance as `TRUE`, proximity as `FALSE`, number of trees to `1000`, tcControl as our above specified `tc` on the training data. (This may take some time, so you might want to start from a small number of trees and proceed with it. Once the codes are done, return to this question and set `ntree=1000`.)

```{r}
x <- data.train %>% select(!year) %>% select(!warstds)
y <- as.factor(data.train$warstds)

tree_mod <- train(x, y, metric = "ROC", sampsize = c(30,90), importance = TRUE, proximity = FALSE, ntrees = 10, trControl = tc)

```


What are the types of variables that seem to feature most in each type of model as predictors?

```{r}

```


Save all models (total $3 \times 2 + 1 = 7$ models) with easy to read names.

```{r}

# Defining list of the 7 models
models <- list(FL_mod, FL_mod_plr, HS_mod, HS_mod_plr, CH_mod, CH_mod_plr)

```



## Question 2

We will now create ROC plots for different models:

- Collect the predicted probabilities for the outcome from each of the above models (note these should be for the highest AUC score in the caret CV procedure, `your-logit-model$finalModel$fitted.values`). For the random forests model, requires a call from `predict()` with type set to "prob" (i.e., `predict(your-rf-model$finalModel, type="prob")`).
- Follow the sample code below to create a `prediction` object from which to calculate the performance of the classifier in terms of true positive and false positive rates.
- Plot the ROC curves of all the unpenalized models (= classic logistic regression) and the RF model.
- Then separate, plot the ROC curves of all penalized models and the RF model. How does the RF model compare?

Sample code: 
```{r eval = FALSE}
## ROC plot: Example with a *classic logistic regression* model trained with caret CV procedure
library(ROCR) # We will use prediction() & performance() functions from this package


results <- c()
for(model in models) {
  
  ## 1. Collect the predicted probabilities
  predict_probas <- model$finalModel$fitted.values 
  # The above line should be changed for *penalized logistic regression* and *RF* model
    
  ## 2. Using in-sample prediction, calculate true positive and false positive rates.
  predictions <- prediction(predict_probas, data.train$warstds)
  results <- c(results, performance(predictions,"tpr","fpr"))

}

# Different predict function for forest
tree_probas <- tree_mod$finalModel$fitted.values 
tree_predicts <- predict(tree_mod$finalModel, type="prob")
results <- c(results, performance(tree_predicts,"tpr","fpr"))

  



```

## Question 3

Finally, we will evaluate out of sample prediction of unpenalized models and the RF model.

Pull out the testing data (1999, 2000) and evaluate each of the model predictions on the testing data. Focus on true positives, or when `warstds` is (correctly) classified with high probability as a "war".
